Density-dependent selection and the limits of relative fitness

### 2018-02-22 manuscript.tex
l

The constant-$N$ behavior of $c$-selection arises from the role of $c$ as a trait determining relative competitive success in territorial contests. As such, this behavior is a result of a reproductive excesses Eq.~\eqref{eq:canonical}. By contrast, previous models of selection-independent density either used unusual models of competition \citep{kimura1969natural,nei1971fertility}, or made restrictive parameter choices in the Lotka-Volterra model (Appendix C; \citealt{smouse_1976,mallet_2012}).

@@

\section*{Appendix A: Poisson approximation}

For simplicity of presentation, we have assumed a Poisson distribution for the $x_i$ as our model of dispersal. Strictly speaking, the total number of $i$ propagules $\sum x_i$ (summed over unoccupied territories) is then no longer a constant $m_i$, but fluctuates between generations for a given mean $m_i$, which is more biologically realistic. Nevertheless, since we do not consider the random fluctuations in type abundances here, and for ease of comparison with the classic lottery model, we ignore the fluctuations in $m_i$. Instead we focus, on Poisson fluctuations in propagule composition in each territory.

@@

The second component, $\Delta_r n_i$, accounts for territories where a single focal propagule is present along with at least one non-focal propagule ($r$ stands for ``rare'') i.e. $x_i=1$ and $X_i\geq 1$ where $X_i=\sum_{j\neq i} x_j$ is the number of nonfocal propagules. The number of territories where this occurs is $Up_i(1)P(X_i\geq 1)=b_i n_i e^{-l_i}(1-e^{-(L-l_i)})$. Thus 
\begin{equation}
\Delta_r n_i = m_i e^{-l_i}(1-e^{-(L-l_i)})\left\langle  \frac{c_i}{c_i +\sum_{j\neq i} c_j x_j } \right\rangle_{\tilde{p}},  \label{eq:deltr}
\end{equation}
where $\langle \rangle_{\tilde{p}}$ denotes the expectation with respect to $\tilde{p}$, and $\tilde{p}$ is the probability distribution of nonfocal propagule abundances $x_j$ \textit{after} dispersal (the significance of evaluating after dispersal should become clearer below), in those territories where exactly one focal propagule, and at least one non-focal propagule, landed.

>>>>

The constant-$N$ behavior of $c$-selection arises from the role of $c$ as a trait determining relative competitive success in territorial contests. As such, this behavior is a result of a reproductive excesses. By contrast, previous models of selection-independent density either used unusual models of competition \citep{kimura1969natural,nei1971fertility}, or made restrictive parameter choices in the Lotka-Volterra model (Appendix C; \citealt{smouse_1976,mallet_2012}).

@@

\section*{Appendix A: Poisson dispersal}

For simplicity of presentation, we assume a Poisson distribution for the $x_i$ as our model of dispersal. Strictly speaking, the total number of $i$ propagules $\sum x_i$ (summed over unoccupied territories) is then no longer a constant $m_i$, but fluctuates between generations for a given mean $m_i$. Nevertheless, since we do not consider the random fluctuations in type abundances here, and for ease of comparison with the classic lottery model, we ignore the fluctuations in $m_i$. Instead we focus on Poisson fluctuations in propagule composition in each territory.

@@

The second component, $\Delta_r n_i$, accounts for territories where a single focal propagule is present along with at least one non-focal propagule ($r$ stands for ``rare'') i.e. $x_i=1$ and $X_i\geq 1$ where $X_i=\sum_{j\neq i} x_j$ is the number of nonfocal propagules. The number of territories where this occurs is $Up_i(1)P(X_i\geq 1)=m_i e^{-l_i}(1-e^{-(L-l_i)})$. Thus 
\begin{equation}
\Delta_r n_i = m_i e^{-l_i}(1-e^{-(L-l_i)})\left\langle  \frac{c_i}{c_i +\sum_{j\neq i} c_j x_j } \right\rangle_{\tilde{p}},  \label{eq:deltr}
\end{equation}
where $\langle \rangle_{\tilde{p}}$ denotes the expectation with respect to $\tilde{p}$, and $\tilde{p}$ is the probability distribution of nonfocal propagule abundances $x_j$ \textit{after} dispersal (the significance of evaluating after dispersal should become clearer below), in those territories where exactly one focal propagule, and at least one non-focal propagule, landed.
